---
title: "Matching simulations 2 - common support for uniform distributions"
author: "Laura Jansen-Storbacka"
date: "4/21/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
source("matching_simulation_functions.R")
library(ggplot2)
library(MatchIt)
library(cobalt)
library("lmtest")
library("sandwich")
library("boot")

```

# Varying Overlap (common support) for 2 uniform distributions

### simulation 1: varying common support/overlap for 3 methods.

The treated group is fixed at unif (0.5,1). the control group slides from unif(0.5,1) to unif(0,0.5), it is always 0.5 units

```{r common-support-n50, cache=TRUE, echo=FALSE}

set.seed(18)
reps <- 1000
n=50
coverage = seq(0,0.5,0.02)

m.1.1 <-mapply(g.x,m=reps,n=n,min_C = 0+coverage, max_C = 0.5+coverage, min_T = 0.5, max_T = 1, replace = FALSE)
dim(m.1.1) <- c(3,reps,length(coverage))                  # change dimensions from list to 3d array

est.1.1 <-apply(m.1.1,c(1,3),mean)[1,]
bias.1.1 <- est.1.1-1                                #vector of biases for each sample size
var.1.1 <- apply(m.1.1,c(1,3),var)[1,]               #vector of variances for each n
mse.1.1 <- var.1.1 + bias.1.1^2

prop_cont.1.1 <- apply(m.1.1,c(1,3),mean)[2,]/n      # average proportion of controls used - always n here
discard_treat.1.1 <- apply(m.1.1,c(1,3),mean)[3,]/n  # average proportion discarded from treatment


```

```{r common_support_unif_replace, cache=TRUE, echo=FALSE}
### Simulation 1.2 varying common support overlap, with replacement
set.seed(18)
reps <- 1000
n=50
coverage = seq(0,0.5,0.02)

m.1.2 <-mapply(g.x,m=reps,n=n,min_C = 0+ coverage, max_C = 0.5+coverage, min_T = 0.5, max_T = 1, replace = TRUE)
dim(m.1.2) <- c(3,reps,length(coverage))                  # change dimensions from list to 3d array

est.1.2 <-apply(m.1.2,c(1,3),mean)[1,]
bias.1.2 <- est.1.2 -1                                #vector of biases for each sample size
var.1.2 <- apply(m.1.2,c(1,3),var)[1,]               #vector of variances for each n
mse.1.2 <- var.1.2 + bias.1.2^2

prop_cont.1.2 <- apply(m.1.2,c(1,3),mean)[2,]/n      # average proportion of controls used - always n here
discard_treat.1.2 <- apply(m.1.2,c(1,3),mean)[3,]/n  # average proportion discarded from treatment

```

```{r common-support-calipers-unif, cache=TRUE, echo=FALSE}
### Simulation 1.3 varying common support overlap, caliper = 0.3
set.seed(18)
reps <- 1000
n=50
coverage = seq(0.16,0.5,0.02)

m.1.3 <-mapply(g.x,m=reps,n=n,min_C = 0 + coverage, max_C = 0.5 + coverage, 
               min_T = 0.5, max_T = 1, replace = FALSE,
               distance = "logit",caliper = 0.3 )

dim(m.1.3) <- c(3,reps,length(coverage))                  # change dimensions from list to 3d array

est.1.3 <-apply(m.1.3,c(1,3),mean)[1,]
bias.1.3 <- est.1.3 -1                                #vector of biases for each sample size
var.1.3 <- apply(m.1.3,c(1,3),var)[1,]               #vector of variances for each n
mse.1.3 <- var.1.3 + bias.1.3^2

prop_cont.1.3 <- apply(m.1.3,c(1,3),mean)[2,]/n      # average proportion of controls used - always n here
discard_treat.1.3 <- apply(m.1.3,c(1,3),mean)[3,]/n  # average proportion discarded from treatment
bias.1.3 <- c(rep(NA,8), bias.1.3)
var.1.3 <- c(rep(NA,8),var.1.3)
mse.1.3 <- c(rep(NA,8),mse.1.3)
discard_treat.1.3 <- c(rep(NA,8), discard_treat.1.3)
prop_cont.1.3 <- c(rep(NA,8), prop_cont.1.3)
```

#### Plotting bias for different overlap levels for uniform data

We see that the caliper = 0.3 give an unbiased estimate, however because of the discarded treatment units, it is not measuring the ATT only the ATM. The caliper method can give an error and not find matches when the overlap is less than around 0.35.

```{r, echo=FALSE}
cover <- seq(0,1,0.04)
plot(cover,bias.1.1, type = "l", col = "steelblue", ylim = c(0,0.8),
     #main = "Bias with different common support proportions", 
     xlab = "Common Support Proportion",
     ylab = "Bias")
lines(cover,bias.1.2, type = "l", col = "red2")
lines(cover,bias.1.3, col="green2")

legend("topright", 
  legend = c("Without replacement", "With replacement", "without replacement, Caliper = 0.3"),
  col = c( "steelblue", "red2", "green2"),lwd=2
  ) 
```

### Plotting variance with different common support proportions, uniform distributions

```{r, echo=FALSE}
cover <- seq(0,1,0.04)
plot(cover,var.1.1, type = "l", col = "steelblue", ylim = c(0,0.27),
     #main = "Variance with different common support, uniform distributions", 
     xlab = "Common Support Proportion",
     ylab = "Variance")
lines(cover,var.1.2, type = "l", col = "red2")
lines(cover,var.1.3, col="green2")

legend("topright", 
  legend = c("Without replacement", "With replacement", "without replacement, Caliper = 0.3"),
  col = c( "steelblue", "red2", "green2"),
  lwd=2) 
```

#### Plotting Mean Squared Error for different overlap levels for uniform data

The MSE is lowest when using a caliper. However as before, because of the hard cut off caused by the uniform distribution, values not in the common support get a propensity score of 0 or 1. This means when the overlap is small the caliper method may not make nay matches and the method breaks down. Additonall, the caliper method uses fewer of the treated units as the overlap becomes smaller, so the matched dataset is smaller and provides less information about the treated group.

```{r, echo=FALSE}
cover <- seq(0,1,0.04)
plot(cover,mse.1.1, type = "l", col = "steelblue", ylim = c(0,0.6),
     #main = "MSE with different common support proportions", 
     xlab = "Common Support Proportion",
     ylab = "Mean Squared Error")
lines(cover,mse.1.2, type = "l", col = "red2")
lines(cover,mse.1.3, col="green2")

legend("topright", 
  legend = c("Without replacement", "With replacement", "without replacement, Caliper = 0.3"),
  col = c( "steelblue", "red2", "green2"),
  lwd=2) 
```

### Plotting the proportion of treated discarded vs overlap when using a caliper

We see that as the proportion of treaded units discarded decreases wth the overlap. With an overlap of 0.6 or less the proportion of discards is over 60% and as the overlap decreases the caliper algorithm may fail to make any matches.

```{r}
cover <- seq(0,1,0.04)


plot(cover, discard_treat.1.3, type = "l", col = "steelblue", ylim = c(0,0.9),lwd=2,
     main = "Proportion discarded using calipers", xlab = "Common Support Proportion",
     ylab = "Proportion treated discarded")
abline(h=seq(0.2,0.8,0.2),lty=2,col="grey")
```

```{r}
cover <- seq(0,1,0.04)


plot(cover, prop_cont.1.3, type = "l", col = "steelblue", ylim = c(0,1),lwd=2,
     main = "Proportion unique controls (With replacement)", xlab = "Common Support Proportion",
     ylab = "Proportion treated discarded")
abline(h=seq(0.2,0.8,0.2),lty=2,col="grey")


```

When sampling with replacement, the proportion of unique controls used increases with overlap. This means that with a smaller overlap, there will be many duplicated control units. This will improve bias but decrease variance.

```{r}

n=10:20


#mean(g.x(n=10,m=50,method="cem",distance=NULL)[1,])

for (i in 10:12){
 print(g.x(n=i,m=10,distance = NULL, method="cem") )
}



```
